### 2014-11-27 main.tex

\documentclass[runningheads,a4paper]{llncs}

\usepackage{style/packages}

\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter  % start of an individual contribution

\title{Data Flow Analysis\\in the Presence of Correlated Calls}

\author{
 Marianna Rapoport\inst{1} \and
 Ond\v{r}ej Lhot\'ak\inst{1} \and
 Frank Tip\inst{2}
}
\institute{
 University of Waterloo\\
 \email{ \{mrapoport, olhotak\}@uwaterloo.ca }
  \and
 Samsung Research America\\
 \email{ ftip@samsung.com  }
}

\pagenumbering{gobble}

\toctitle{Data Flow Analysis in the Presence of Correlated Calls}
%\tocauthor{Authors' Instructions} % TODO
\maketitle

\begin{abstract}
We present a technique to improve the precision of data-flow analyses on object-oriented programs in the presence of 	\textit{correlated calls}. Two method calls are correlated if they are polymorphic and are invoked on the same object. Correlated calls are problematic because they can make  existing data-flow analyses consider certain infeasible data-flow paths as valid. This leads to loss in precision of the analysis solution.

We show how infeasible paths can be eliminated for \textit{Inter-procedural Finite Distributive Subset} (IFDS) problems, a large class of data-flow analysis problems. We show how the precision of IFDS problems can be improved in the presence of correlated calls, by using the \textit{Inter-procedural Distributive Environment} (IDE) algorithm to eliminate infeasible paths.  Using IDE, we eliminate the infeasible paths and obtain a more precise result for the original IFDS problem.

Our analysis is implemented in Scala, using the WALA framework for static program analysis on Java bytecode.

%\keywords{We would like to encourage you to list your keywords within
%the abstract section}
\end{abstract}

\input{style/commands}

\input{content/intro}
\input{content/background}
\input{content/impl}
\input{content/datastr}
\input{content/eval}
\input{content/concl}

\bibliographystyle{plain}
\bibliography{bib/ref}
\end{document}