### 2014-12-17 main.tex

\documentclass[runningheads,a4paper]{llncs}

\usepackage{style/packages}

\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\input{style/commands}

\pagenumbering{gobble}
\listoftodos[Overview of Comments]

\mainmatter  % start of an individual contribution

\title{Precise Data Flow Analysis\\in the Presence of Correlated Method Calls}
\titlerunning{Data Flow Analysis in the Presence of Correlated Calls}

\author{
 Marianna Rapoport\inst{1} \and
 Ond\v{r}ej Lhot\'ak\inst{1} \and
 Frank Tip\inst{2}
}
\institute{
 University of Waterloo\\
 \email{ \{mrapoport, olhotak\}@uwaterloo.ca }
  \and
 Samsung Research America\\
 \email{ ftip@samsung.com  }
}

\pagenumbering{arabic}

\toctitle{Data Flow Analysis in the Presence of Correlated Calls}
%\tocauthor{Authors' Instructions} % TODO
\maketitle

\input{content/abstract}
\input{content/intro}
\input{content/example}
\input{content/background}
\input{content/impl}
%\input{content/datastr}
%\input{content/eval}
\input{content/related}
\input{content/concl}

\bibliographystyle{plain}
\bibliography{bib/ref}

\input{content/appendix}
\end{document}