### 2014-12-17 content/concl.tex
rewrote conclusions

\section{Conclusions}
  \label{sec:Conclusions}

Previous algorithms for dataflow analysis are unable to avoid propagating
dataflow facts along infeasible paths that arise in the presence of
correlated polymorphic method calls. 
% 
We present an approach for transforming an IFDS problem into an IDE problem
in which path feasibility is encoded into the functions that are associated
with edges in an exploded control-flow supergraph. The solution to this
IDE problem can be mapped back to the solution space of the original IFDS
problem, and is potentially more precise because the flow of data along
infeasible paths is prevented. We present a formalization of the
transformation and prove its correctness. We also briefly report on
preliminary experiments with our prototype implementation. Full proof
details are available as supplemental materials with this submission,
and we plan to submit our implementation and experimental data to the
artifact evaluation committee.