### 2014-12-17 content/abstract.tex
spell check

\begin{abstract}
    When two methods are invoked on the same object, the dispatch behaviours of these
    method calls will be correlated. If two correlated method calls are polymorphic 
    (i.e., they dispatch to different method definitions depending on the type of 
    the receiver object), a program's interprocedural control flow graph (ICFG)
    will contain infeasible paths. Existing algorithms for dataflow analysis 
    are unable to ignore such infeasible paths, giving rise to loss of precision.

	We show how infeasible paths due to correlated calls can be eliminated for 
	\textit{Interprocedural Finite Distributive Subset} (IFDS) problems, a large class 
	of data-flow analysis problems with broad applications. Our approach is to
	transform an IFDS problem into an \textit{Interprocedural Distributive 
	Environment} (IDE) problem, in which edge functions are designed in a
	way that enables us to ignore dataflow along infeasible paths. A solution
	to this IDE problem can be mapped back to the solution space of the original
	IFDS problem.
        We formalize the approach, prove it correct, and report on an implementation
        in the WALA analysis framework.
\end{abstract}