### 2015-03-05 ecoop2014/reviews/response.txt
reviews

esponse by Marianna Rapoport <mrapoport@uwaterloo.ca>
   Paper #25: Precise Data Flow Analysis in the Presence of Correlated
              Method Calls
---------------------------------------------------------------------------
Reviewers C and D correctly point out that our primary contributions are
theoretical: the transformation of an IFDS problem to an IDE problem
that provides more precise results for the original IFDS problem, and
the proofs that the result of this approach is both sound and fully
eliminates flow along infeasible paths due to correlated calls.

Reviewers B and C ask about object sensitivity. Object-sensitive
points-to analysis is orthogonal to our work because it analyzes the
flow of data (pointers), while we analyze control flow paths instead.
Also, object-sensitive points-to analysis is flow-insensitive, while
IFDS/IDE are flow-sensitive, and our transformation only makes sense
in a flow-sensitive setting (since a flow-insensitive analysis already
introduces many more infeasible control flow paths).

It would be possible to simulate the effect of our correlated calls
transformation in the following way inspired by object-sensitive
points-to analysis: we could re-analyze each method in a number
of contexts. There would be a separate context for every possible
assignment of concrete types to all of the pointers in the method that
are used as receivers at a call site. The number of such contexts for
each method would be O(R^T), where R is the number of receiver pointers
in the method and T is the number of possible concrete types that could
be assigned to a receiver pointer. Our approach computes equally precise
analysis results but avoids this exponential cost.

Reviewer C says it would be of interest to know how many calls share the
same base variable on average. The average number of polymorphic calls per
receiver is included in Table 1, and is equal to 3. The frequency of
 situations where more than two polymorphic calls are invoked on the same
receiver, compared to all calls in the program, is reflected by the number
of correlated calls over polymorphic calls, which is 39%, and number of
polymorphic calls over all calls, which is 3%.

Reviewer B asks about the inclusion of our IDE implementation in WALA.
We have been in discussion about this with the key WALA developers since
the summer, and we have already begun the process of contributing our
IDE implementation (and correlated calls transformation) to WALA so that
it will be available to all WALA users.

We agree with reviewer D that the asymptotic complexity is O(ED^3RT) and will present it more explicitly in the final version.

Reviewer A asks why we do not just use IDE directly. The IFDS and
IDE algorithms solve two distinct classes of problems. Our approach
is aimed at IFDS problems rather than IDE problems. Given an IFDS
problem, our approach results in an IFDS solution that is more precise
than if the same IFDS problem had been solved directly with the IFDS
algorithm. As future work, we could look at adapting our approach to IDE
problems: we could take an initial IDE problem, convert it to a larger
and more complex IDE problem, so that the solution of the larger IDE
problem generates a more precise solution to the original IDE problem by
removing infeasible paths due to correlated calls.