### 2015-03-09 sas2015/content/concl.tex
squishing (mainly) of introduction

\section{Conclusions}
  \label{sec:Conclusions}

Previous algorithms for data-flow analysis are unable to avoid propagating
data-flow facts along infeasible paths that arise in the presence of
correlated polymorphic method calls. 
% 
We present an approach for transforming an IFDS problem into an IDE problem
in which path feasibility is encoded into the functions that are associated
with edges in an exploded control-flow supergraph. The solution to this
IDE problem can be mapped back to the solution space of the original IFDS
problem, and is more precise for some client programs because the flow of data along
infeasible paths is prevented. We present a formalization of the
transformation and prove its correctness. We also briefly report on
preliminary experiments with our prototype implementation. Full proof
details are available in the Appendix,
and we plan to submit our implementation and experimental data to the
artifact evaluation committee.

As future work, it is possible to adapt our approach to work on IDE
problems in addition to IFDS problems. We would take an initial IDE
problem, convert it to a larger and more complex IDE problem, such
that the solution of the larger IDE problem generates a more precise
solution to the original IDE problem by removing infeasible paths due to
correlated calls.