### 2015-03-09 sas2015/content/intro.tex

\section{Introduction}

\commentout{
Data-flow analysis computes an approximation of how values may flow  
through a program, and has applications in compiler optimization, 
programming tools, and computer security, and many other areas.
Data-flow analyses operate on \textit{control-flow graphs} (CFGs)
that model 
the order in which the instructions of a program are executed. This is
typically done by associating \textit{flow functions} that represent how 
data is propagated with the edges of the CFG.  A 
\textit{confluence operator} specifies how the data facts that have been 
computed along different paths should be merged when the paths join.
}

%Since a CFG is an over-approximation of the possible flows 
A control-flow graph (CFG) is an over-approximation of the possible flows 
of control in concrete executions of a program. It may contain \textit{infeasible} 
paths that cannot occur at runtime. The precision of a data-flow analysis algorithm
depends on its ability to detect and disregard such infeasible paths.
The \textit{Interprocedural Finite Distributive Subset} (IFDS) algorithm~\cite{reps1995precise}
is one general data-flow analysis algorithm that 
%by  Reps, Horwitz, and Sagiv
%is a general data-flow analysis
%algorithm
%for computing solutions to finite distributive data-flow problems such 
%that can be instantiated to solve problems such
%as reaching definitions, available expressions, and taint analysis.
%A distinguishing 
%characteristic of IFDS is that it
avoids infeasible interprocedural paths in which calls 
and returns to/from functions are not properly matched. %Sagiv, Reps, and Horwitz
%also presented
The \textit{Interprocedural Distributive Environment} (IDE) algorithm~\cite{sagiv1996precise}
has the same property, but
%that similarly only considers properly matched call/return
%edges, but that 
supports a broader range of data-flow problems.
%by expanding the 
%domain of flow functions to \textit{environments} that go beyond the data-flow facts 
%considered by IFDS.

This paper presents an approach to data-flow analysis that avoids a type
of infeasible path that arises in object-oriented programs when two or more
methods are dynamically dispatched on the same receiver object.
%\mtodo{Should we define the term ``receiver''?}
%In such cases,
If the
method calls are polymorphic
(i.e. the method invoked depends on the run-time type of the receiver),
%(i.e., if they dispatch to different method definitions 
%depending on the type of the receiver expression at run time),
then their dispatch 
behaviours are correlated, and some of the paths between them are infeasible. A recent paper~\cite{DBLP:journals/scp/Tip15} made this observation
but did not present any concrete algorithm to take advantage of it.
%, and we are not aware 
%of any existing data-flow analysis that is capable of avoiding infeasible paths that arise 
%in the presence of correlated method calls.
% (see Section~\ref{sec:Related} for a discussion
%of related work).

%\input{figures/fig-transformation}

%The approach taken in our work is to 
Our approach transforms an IFDS problem into an
IDE problem that precisely accounts for infeasible paths due to correlated calls. 
The results of this IDE problem can be mapped back to the data-flow domain of the 
original IFDS problem, but are more precise than the results of directly applying
the IFDS algorithm to the original problem.
%
We present a formalization of the transformation and prove its correctness:
specifically, we prove it still soundly considers all paths that are feasible,
and that it avoids flow along all paths that are infeasible due to
correlated calls.
%as follows. 
%First, we derive an IDE problem that is equivalent to the original IFDS problem,
%as described by Reps et al.~\cite{sagiv1996precise}.
%Solving this ``Equivalence-IDE Problem'' with the IDE algorithm yields a result that
%is equivalent to the one obtained by applying the IFDS algorithm to the original IFDS problem.
%Then, we derive a ``Correlated-Calls 
%IDE Problem'' from the original IFDS problem and show 
%that the solution to this problem, when converted to an IFDS result, is more precise
%than the original IFDS result.
%We also show that the correlated-calls analysis is sound, i.e., that it never considers concrete execution paths as infeasible.
%This is illustrated schematically
%in Figure~\ref{fig:transformations}%
%\footnote{ 
%  The labels %$\resultIDE$, $\resultIFDS$, $\backEq$, $\backCC$, $\transEq$, and $\transCC$
%  on edges in 
%  Figure~\ref{fig:transformations} reflect a number of mappings and %projections 
%  that will be defined in Section~\ref{sec:cca} and that can be %ignored here.
%}.

We implemented the correlated-calls transformation and the IDE algorithm in Scala,
on top of the WALA framework for static analysis of JVM bytecode~\cite{fink2012wala}.
Our prototype implementation was tested extensively by using it to transform an IFDS-based 
taint analysis into a more precise IDE-based taint analysis, and applying the latter
to small example programs with correlated calls. Our prototype along with all tests
will be made available to the artifact evaluation committee.

%We also report on preliminary experiments in which our correlated-calls transformation
%is applied to an IFDS formulation of a simple taint analysis. Our results show that
%solving the resulting IDE problem avoids infeasible paths due to correlated calls as
%expected.
%
%In summary, the contributions of this paper are as follows:
%\begin{itemize}
%  \item
%    We present a general approach for transforming IFDS problems into corresponding
%    IDE problems that avoid infeasible paths due to correlated method calls and
%    prove its correctness. 
%  \item
%    We implemented the approach in Scala, on top of the WALA program analysis framework
%    and report on preliminary experiments. 
%\end{itemize}

The remainder of this paper is organized as follows.
%
Section~\ref{sec:MotivatingExample} presents a motivating example.
%
Section~\ref{sec:bg} reviews the IFDS and IDE algorithms.
%
Section~\ref{sec:cca} presents the correlated-calls transformation
and sketches a proof of its correctness%
\footnote{
  Detailed proofs of our lemmas and theorems can be found in
  the appendix.
}.
%
Related work is discussed in Section~\ref{sec:Related}.
%
Finally, conclusions and directions for future work are presented in Section~\ref{sec:Conclusions}.