### 2015-06-19 sas2015_final/main.tex

\usepackage{style/packages}

\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\input{style/commands}

\pagenumbering{gobble}
\listoftodos[Overview of Comments]

\mainmatter

\title{Precise Data Flow Analysis\\in the Presence of Correlated Method Calls\footnote{This research was supported by the Natural Sciences and Engineering Research Council of Canada and the Ontario Ministry of Research and Innovation.}}
\titlerunning{Data Flow Analysis in the Presence of Correlated Calls}
\techreport{\subtitle{David R. Cheriton School of Computer Science Technical Report CS-2015-07}}

\author{
 Marianna Rapoport\inst{1} \and
 Ond\v{r}ej Lhot\'ak\inst{1} \and
 Frank Tip\inst{2}
}
\institute{
 University of Waterloo\\
 \email{ \{mrapoport, olhotak\}@uwaterloo.ca }
  \and
 Samsung Research America\\
 \email{ ftip@samsung.com  }
}

\pagenumbering{arabic}

\toctitle{Data Flow Analysis in the Presence of Correlated Calls}

\maketitle

\input{content/abstract}
\input{content/intro}
\input{content/example}
\input{content/background}
\input{content/impl}
\input{content/related}
\input{content/concl}

\vspace{.8cm}
\makeatletter
\patchcmd{\@makeschapterhead}{\vskip40}{\vskip10}{}{}
\makeatother

\bibliographystyle{myplain}
\begingroup
\let\clearpage\relax
\bibliography{bib/ref}
\endgroup

\techreport{\newpage\input{content/appendix.tex}}
\end{document}